Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-27 article.tex
conclusions

% Possible extensions
% partial autocorrelation function (sometimes "partial correlation function")
% conditional independence test
% Speak about the complexity and parallelization of the methods

In this paper, we present the best performing method of the Connectomics challenge. Our
approach is simple and theoretically sounded and made of two main stages: the
signal is first preprocessed with several filters and then we derive the degree
of association matrix by computing the partial correlation statistics matrix.
We also evaluate the method in a more detailed way and perform a comparison
with other inference methods. We finally discuss the relevance of the metric
used in the challenge and we point out that it was not necessary to discover
causal relationships to maximize the ROC AUC. Thus, it may be  promising to
focus on that particular topic in future works.

% In the discussion, we give hints about limiting the depth of trees in the
% GENIE3 method to infer the network. It should be interesting to see if there
% is a optimum value for the maximal depth. And of course, it may also rewarding
% to increase considerably the number of trees to get more stable predictions.

% In previous study \citep{shipley2002cause}, it has been suggested to
% consider correlation coefficients of all possible orders,
% i.e. conditioned on every subsets of the set of $p-2$ other variables. This
% would take into account multiple indirect paths between a given pair of
% variables. Instead of considering only the highest order, it may be useful to
% study with accuracy the inferred network at all orders. In this case, it may not
% be very relevant but interesting in case of dynamical networks.

% As we mentioned, in order to maximize the ROC AUC, it was not necessary to
% discover causal relationships. It may be, in future works, promising to focus
% on that particular topic.

\section*{Acknowledgments}
A Joly and G Louppe are research fellows of the FNRS, Belgium.  A Sutera is
recipient of a FRIA fellowship of FRS-FNRS, Belgium. This work is supported by
PASCAL2 and the IUAP DYSCO, initiated by the Belgian State, Science Policy
Office.

>>>>

In this paper, we outlined a simple but efficient methodology for the problem
of connectome inference from calcium imaging data. Our approach is made of two
steps: (i) processing fluorescence data to detect neural peak activities and
(ii) inferring the degree of association between neurons from partial
correlation statistics. Its simplified variant shows to outperform other
network inference methods while its optimized version proved to be best method
on the Connectomics challenge. Given its simplicity and good performance, we
therefore believe that the (simplified) methodology presented in this work
would constitute a solid and easily reproducible baseline for further work in
the field of connectome inference.

\section*{Acknowledgments}
A. Joly and G. Louppe are research fellows of the FNRS, Belgium.  A. Sutera is
recipient of a FRIA fellowship of FRS-FNRS, Belgium. This work is supported by
PASCAL2 and the IUAP DYSCO, initiated by the Belgian State, Science Policy
Office.
